---
title: "ANON CODE"
author: "Francisco Benitez and Anelisse DOminicci"
date: "Replicathon 2019"
output:
  html_document: default
  pdf_document: default
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
```

## Instructions:

In this R Markdown document, you and your team will create a fully reproducible analysis with the goal of assessing and interpreting the replicability of two pharmacogenomic experiments. This document should contain all of the text and code of your analyses, which will allow others to run, interpret, and reuse your work.

The questions below will help guide you in your analyses and interpretation of results. You don't need to answer every question, but for the problems you do complete, make sure that you completely justify your conclusions by explaining your reasoning and including numerical summaries and data visualizations wherever possible. There are four tutorials (also R Markdown documents) that will help you learn new tools to tackle these problems, and the questions are divided into four sections corresponding to the tutorials (though many overlap with more than one tutorial). If questions arise during your analyses that do not fit into these problems, feel free to include those as well. 

For each answer, include text by simply typing below the question. Include code in code blocks (include three back ticks at the start and end of each code block):

```{r}
#Your code goes here
```

You may find it helpful to use the version control and code sharing system called GitHub to work together with your team so that all of you can edit the same document and keep track of its changes. Here is a [setup guide](https://github.com/datasciencelabs/2016/blob/master/lectures/git-and-github/setting-up-git.Rmd) and [brief introduction to Git and GitHub](https://github.com/datasciencelabs/2016/blob/master/lectures/git-and-github/version-control.Rmd) from another course. The mentors will be able to help if you run into problems. 

## Questions:

### Exploratory analysis of pharmacogenomic data

1. How many cell-lines are contained in the data?
```{r}
rawPharmacoData <- readRDS("C:/Users/HP/Desktop/PR2019replicathon/data/rawPharmacoData.rds")
unique(rawPharmacoData$cellLine)
```
```{r}
Celllines <- unique(rawPharmacoData$cellLine) 
length(Celllines)
```

2. What drug concentrations were used in each study?
```{r}
Drug <- unique(rawPharmacoData$drug) 
length(Drug)
```

```{r}
rawPharmacoData %>%
    group_by(study) %>%
    summarize(Concentration = n_distinct(concentration))
```
```{r}
rawPharmacoData %>%
    group_by(study) %>%
    summarize(Concentration = n_distinct(concentration), droga = n_distinct(drug))
```
```{r}
rawPharmacoData %>%
  group_by(study)%>%
  count(concentration)
```



3. Histograms, appart from telling us how the data is
distributed, can also make evident potential
problems with the data. Plot a histogram
of drug viabilities. Does it look as one would
expect from the description of the data?

```{r conc-hist3, fig.width = 8, fig.height = 4, warning = FALSE}
rawPharmacoData %>%
    ggplot(aes(x = log2(viability))) +
    geom_histogram(fill = "gray", color = "black") +
    facet_wrap(~ drug) +
    ggtitle("Distributions of viability by drug")
```

```{r conc-hist2, fig.width = 8, fig.height = 4, warning = FALSE}
rawPharmacoData %>%
    ggplot(aes(x = viability)) +
    geom_histogram(aes( color = study), binwidth = 0.1) +
    facet_wrap(~ drug) +
    ggtitle("Distributions of viability by drug")
```


4. How many viability data points are within the
expected range according to the definition of viability (e.g. above 0 
and below 100)? Try to come up with explanations about
the values that are out of range. Are these due to technical
issues? Biology?

```{r}
rawPharmacoData %>%
    group_by(study)%>%
    summarize(min_viability = min(viability),
              max_viability = max(viability),
              n_too_small   = sum(viability < 0),
              n_too_big     = sum(viability > 100),
              n_middle      = sum(viability >0 & viability < 100))
```


5. Read the csv file containing the summarized data. What kinds
of variables are in the data? What does each column represent

```{r}
summarizedPharmacoData <-readRDS("C:/Users/HP/Desktop/PR2019replicathon/data/summarizedPharmacoData.rds")
str(summarizedPharmacoData)
```





6. Plot histograms of the viability scores at different levels of drug
doses. Are stronger drug concentrations consistent with lower viability scores?

-No necesariamente mira la numero 10

```{r conc-hist, fig.width = 8, fig.height = 4, warning = FALSE}
rawPharmacoData %>%
    ggplot(aes(x = viability)) +
    geom_histogram(aes( color = drug), binwidth = 0.1) +
    facet_wrap(~ concentration) +
    ggtitle("Distributions of viability by different concentrations of drugs")
```

### Using Correlation Measures to Assess Replicability of Drug Response Studies

1. Create a scatterplot for each drug comparing the AUC in GDSC and CCLE for all cell lines (hint: code from Tutorial 1b may help).


```{r}
ggplot(summarizedPharmacoData, aes(x = auc_GDSC, y = auc_CCLE, col = drug)) +
    geom_jitter(alpha = 1/2) +
    xlab("GDSC AUC") +
    ylab("CCLE AUC") +
    ggtitle("Comparing AUC in GDSC and CCLE for all cell lines")
```
```{r}
ggplot(summarizedPharmacoData, aes(x = auc_GDSC, y = auc_CCLE, col = drug)) +
    geom_jitter(alpha = 1/2) +
    xlab("GDSC AUC") +
    ylab("CCLE AUC") +
    facet_wrap(~ drug) +
    geom_smooth(method = "lm", col = "black") +
    ggtitle("Comparing AUC in GDSC and CCLE for all drugs")
```
2. Calculate correlation coefficients of the AUC in GDSC and CCLE for each drug (hint: code from Tutorial 1b may help). 

```{r}
drugCorrs <- summarizedPharmacoData%>% 
    group_by(drug) %>%
    summarize(Pearson_ic50  = cor(-log10(ic50_GDSC / 10^6), -log10(ic50_CCLE / 10^6), method = "pearson"),
              Spearman_ic50 = cor(-log10(ic50_GDSC / 10^6), -log10(ic50_CCLE / 10^6), method = "spearman"))

drugCorrs
```

```{}
drugCorrs %>%
  tidyr::spread(measure, correlation) %>%
  ggplot(aes(x = Pearson_ic50, y = Spearman_ic50, label = drug)) +
    geom_point(alpha = 1/2) +
    geom_text() +
    ggtitle("Correlation of cell line IC50 summaries between studies for each drug")
```
drugCorrs <- gather(drugCorrs, measure, correlation, -drug)

drugCorrs
```{r}
drugCorrs <- gather(drugCorrs, measure, correlation, -drug)

drugCorrs
```
```{r barplot-correlations-ic50, fig.width = 8, fig.height = 4}
drugCorrs %>%
    ggplot(aes(x = drug, y = correlation, fill = measure, group = measure)) +
    geom_bar(stat = "identity", position = position_dodge(), colour = "black") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_grey() +
    ylim(0, 1) + 
    ggtitle("Correlation of cell line IC50 summaries between studies for each drug")
```

